---
title: "FORMATIVE ASSESSMENT 4"
author: "Lindsy Rossel Masicat"
date: "2024-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(moments)
library(knitr)
options(scipen = 999) 
```
<style>
  table, caption {
    margin-left: auto;
    margin-right: auto;
    text-align: center;
  }
  caption{
    font-size: 14pt;
    font-weight: bold;
    color: black;
}
</style>
```{r table, echo=FALSE}
normal_data <- c(67, 70, 63, 65, 68, 60, 70, 64, 69, 61, 66, 65, 71, 62, 66, 68, 64, 67, 62, 66, 65, 63, 66, 65, 63, 69, 62, 67, 59, 66, 65, 63, 65, 60, 67, 64, 68, 61, 69, 65, 62, 67, 70, 64, 63, 68, 64, 65, 61, 66)
skewed_right_data <- c(31, 43, 30, 30, 38, 26, 29, 55, 46, 26, 29, 57, 34, 34, 36, 40, 28, 26, 66, 63, 30, 33, 24, 35, 34, 40, 24, 29, 24, 27, 35, 33, 75, 38, 34, 85, 29, 40, 41, 35, 26, 34, 19, 23, 28, 26, 31, 25, 22, 28)
skewed_left_data <- c(102, 55, 70, 95, 73, 79, 60, 73, 89, 85, 72, 92, 76, 93, 76, 97, 10, 70, 85, 25, 83, 58, 10, 92, 82, 87, 104, 75, 80, 66, 93, 90, 84, 73, 98, 79, 35, 71, 90, 71, 63, 58, 82, 72, 93, 44, 65, 77, 81, 77)
uniform_data <- c(12.1, 12.1, 12.4, 12.1, 12.1, 12.2, 12.2, 12.2, 11.9, 12.2, 12.3, 12.3, 11.7, 12.3, 12.3, 12.4, 12.4, 12.1, 12.4, 12.4, 12.5, 11.8, 12.5, 12.5, 12.5, 11.6, 11.6, 12, 11.6, 11.6, 11.7, 12.3, 11.7, 11.7, 11.7, 11.8, 12.5, 11.8, 11.8, 11.8, 11.9, 11.9, 11.9, 12.2, 11.9, 12, 11.9, 12, 12, 12)
```

```{r tableputput, echo=FALSE}
data_table <- data.frame(
  Normal = normal_data,
  Skewed_Right = skewed_right_data,
  Skewed_Left = skewed_left_data,
  Uniform = uniform_data
)
kable(data_table, caption = "Table 5.1", align = c("c", "c", "c", "c"))
```
The data in Table 5.1 samples of size 50 from a normal distribution, a skewed-right distribution, a skewed-left distribution, and a uniform distribution.

The normal data are female height measurements, the skewed-right data are age at marriage for females, the skewed-left data are obituary data that give the age at death for females, and the uniform data are the amount of cola put into a 12 ounce container by a soft drinks machine.

# Compute the following:

## 1. Find the (a) first, (b) second, (c) third, and (d ) fourth moments for each of the sets of data (normal, skewed-right, skewed-left, uniform).

### Normal

```{r normal, echo=FALSE}
m1Normal <- round(sum(normal_data)/length(normal_data),2)
m2Normal <- round(sum(normal_data^2)/length(normal_data),2)
m3Normal <- round(sum(normal_data^3)/length(normal_data),2)
m4Normal <- round(sum(normal_data^4)/length(normal_data),2)
```
(a) First Moment: `r m1Normal`
(b) Second Moment: `r m2Normal`
(c) Third Moment: `r m3Normal` 
(d) Fourth Moment:`r m4Normal`

### Skewed-right
```{r skewed-rigt, echo=FALSE}
m1SkewedRight <- round(sum(skewed_right_data)/length(skewed_right_data),2)
m2SkewedRight <- round(sum(skewed_right_data^2)/length(skewed_right_data),2)
m3SkewedRight <- round(sum(skewed_right_data^3)/length(skewed_right_data),2)
m4SkewedRight <- round(sum(skewed_right_data^4)/length(skewed_right_data),2)
```
(a) First Moment: `r m1SkewedRight`
(b) Second Moment: `r m2SkewedRight`
(c) Third Moment: `r m3SkewedRight` 
(d) Fourth Moment:`r m4SkewedRight`

### Skewed-left
```{r skewed-left, echo=FALSE}
m1SkewedLeft <- round(sum(skewed_left_data)/length(skewed_left_data),2)
m2SkewedLeft <- round(sum(skewed_left_data^2)/length(skewed_left_data),2)
m3SkewedLeft <- round(sum(skewed_left_data^3)/length(skewed_left_data),2)
m4SkewedLeft <- round(sum(skewed_left_data^4)/length(skewed_left_data),2)
```
(a) First Moment: `r m1SkewedLeft`
(b) Second Moment: `r m2SkewedLeft`
(c) Third Moment: `r m3SkewedLeft` 
(d) Fourth Moment:`r m4SkewedLeft`

### Uniform
```{r uniform, echo=FALSE}
m1Uniform <- round(sum(uniform_data)/length(uniform_data),2)
m2Uniform <- round(sum(uniform_data^2)/length(uniform_data),2)
m3Uniform <- round(sum(uniform_data^3)/length(uniform_data),2)
m4Uniform <- round(sum(uniform_data^4)/length(uniform_data),2)
```
(a) First Moment: `r m1Uniform`
(b) Second Moment: `r m2Uniform`
(c) Third Moment: `r m3Uniform` 
(d) Fourth Moment:`r m4Uniform`

## 2. Find the (a) first, (b) second, (c) third, and (d ) fourth moments about the mean for each of the sets of data (normal, skewed-right, skewed-left, uniform).

### Normal

```{r normal2, echo=FALSE}
m1Normal2 <- round(sum(normal_data - m1Normal)/length(normal_data),2)
m2Normal2 <- round(sum((normal_data - m1Normal)^2)/length(normal_data),2)
m3Normal2 <- round(sum((normal_data - m1Normal)^3)/length(normal_data),2)
m4Normal2 <- round(sum((normal_data - m1Normal)^4)/length(normal_data),2)
```
(a) First Moment: `r m1Normal2`
(b) Second Moment: `r m2Normal2`
(c) Third Moment: `r m3Normal2` 
(d) Fourth Moment:`r m4Normal2`

### Skewed-right
```{r skewedright2, echo=FALSE}
m1SkewedRIght2 <- round(sum(skewed_right_data - m1SkewedRight)/length(skewed_right_data),2)
m2SkewedRIght2 <- round(sum((skewed_right_data - m1SkewedRight)^2)/length(skewed_right_data),2)
m3SkewedRIght2 <- round(sum((skewed_right_data - m1SkewedRight)^3)/length(skewed_right_data),2)
m4SkewedRIght2 <- round(sum((skewed_right_data - m1SkewedRight)^4)/length(skewed_right_data),2)
```
(a) First Moment: `r m1SkewedRIght2`
(b) Second Moment: `r m2SkewedRIght2`
(c) Third Moment: `r m3SkewedRIght2` 
(d) Fourth Moment:`r m4SkewedRIght2`

### Skewed-Left
```{r skewedleft2, echo=FALSE}
m1SkewedLeft2 <- round(sum(skewed_left_data - m1SkewedLeft)/length(skewed_left_data),2)
m2SkewedLeft2 <- round(sum((skewed_left_data - m1SkewedLeft)^2)/length(skewed_left_data),2)
m3SkewedLeft2 <- round(sum((skewed_left_data - m1SkewedLeft)^3)/length(skewed_left_data),2)
m4SkewedLeft2 <- round(sum((skewed_left_data - m1SkewedLeft)^4)/length(skewed_left_data),2)
```
(a) First Moment: `r m1SkewedLeft2`
(b) Second Moment: `r m2SkewedLeft2`
(c) Third Moment: `r m3SkewedLeft2` 
(d) Fourth Moment:`r m4SkewedLeft2`

### Uniform

```{r uniform2, echo=FALSE}
m1Uniform2 <- round(sum(uniform_data - m1Uniform)/length(uniform_data),2)
m2Uniform2 <- round(sum((uniform_data - m1Uniform)^2)/length(uniform_data),2)
m3Uniform2 <- round(sum((uniform_data - m1Uniform)^3)/length(uniform_data),2)
m4Uniform2 <- round(sum((uniform_data - m1Uniform)^4)/length(uniform_data),2)
```
(a) First Moment: `r m1Uniform2`
(b) Second Moment: `r m2Uniform2`
(c) Third Moment: `r m3Uniform2` 
(d) Fourth Moment:`r m4Uniform2`

## 3. Find the (a) first, (b) second, (c) third, and (d ) fourth moments about the number 75 for the set of female height measurements.

### Female height measurements

```{r female, echo=FALSE}
m1Female <- round(sum(normal_data - 75)/length(normal_data),2)
m2Female <- round(sum((normal_data - 75)^2)/length(normal_data),2)
m3Female <- round(sum((normal_data - 75)^3)/length(normal_data),2)
m4Female <- round(sum((normal_data - 75)^4)/length(normal_data),2)
```
(a) First Moment: `r m1Female`
(b) Second Moment: `r m2Female`
(c) Third Moment: `r m3Female` 
(d) Fourth Moment:`r m4Female`

### 4. Using the results of items 2 and 3 for the set of female height measurements, verify the relations between the moments (a) $m_2 = m'_2 - m'^2_1$, <or> (b) $m_3 = m'_3 - 3m'_1m'_2 +2m'^3_1$, <or> (c) $m_4 = m'_4 - 4m'_1m'_3 + 6m'^2_1m'_2 - 3m'^4_1$.

```{r verify, echo=FALSE}
message <- "The result is the same. Thus it is correct."

verify2 <- round(m2Female - m1Female^2,2)
m2Verification <- ifelse(m2Normal2 == verify2, message, "The result is different.")

verify3 <- round(m3Female - 3 * m1Female * m2Female + 2 * m1Female^3,2)
m3Verification <- ifelse(m3Normal2 == verify3, message, "The result is different.")

verify4 <- round(m4Female - 4 * m1Female * m3Female + 6 * m1Female^2 * m2Female - 3 * m1Female^4,2)
m4Verification <- ifelse(m4Normal2 == verify4, message, "The result is different.")

```
(a) $m_2 = m'_2 - m'^2_1$
<br>
`r m2Normal2` = `r verify2`
<br>
Verification: `r m2Verification`
<br>
(b) $m_3 = m'_3 - 3m'_1m'_2 +2m'^3_1$
<br>
`r m3Normal2` = `r verify3`
<br>
Verification: `r m3Verification`
<br>
(c) $m_4 = m'_4 - 4m'_1m'_3 + 6m'^2_1m'_2 - 3m'^4_1$
<br>
`r m4Normal2` = `r verify4` 
<br>
Verification: `r m4Verification`